### 2015-07-01 narrative/naming-service.tex
Lots more details. Need to resolve refs still.

A traditional POSIX naming service offers a hierarchical space consisting of
directories and files. LWFS and Sirocco have abandonded this as the default
model in favor of a container/object/fork/addrress tuple for identifying data.
LWFS demonstrated a POSIX-style namespace on the side kept in sync using a
transaction process like D2T~\cite{d2t} showing that this alternative approach
can support traditional POSIX API calls even though the underlying storage
system uses a different model.

For this project, we seek to offer a similar POSIX-style namespace for naming
objects, but store a rich metadata collection with each entry to handle data
migration, discovery, and writing and retrevial time estimations. Further, we
will also have to track data stored in different tiers, even at the variable
subset level. This additional metadata load will require a more complex naming
service that reacts to automatic data movement while caching current data
locations accelerating data access operations avoiding any data discovery
operations.

In addition to the basic naming operations, we will also need to incorporate
authorization capabilities. Sirocco currently integrates with a Kerberos
service for authentication and authorization. Given a capability ticket, a
user can access different objects as needed.

The main challenge of incorporating the additional, rich metadata will be
joined by the additional challenge of coordinating with the other storage
and application-layer services to offer the best access times possible for data
stored in the system.

>>>>

A traditional POSIX naming service offers a hierarchical space consisting of
directories and files. This structure has led to scalability problems because
of the serialized access to a single source for creating and accessing files.
Several efforts~\cite{giga+,pvfs,others} have worked to reduce this contention
by doing things like reducing the serialized scope to a single directory or
subtree. While these approaches help, they do not address this key scalability
limitation.

Pure object stores, such as those popular in the big data
domain~\cite{memcached,others} avoid this bottleneck by strictly offering an
object ID with the application required to manage how this ID maps to something
of interest. This approach of removing the metadata service from the system
level completely can work well for scale out applications where data is created
or consumed by a single process at a time rather than potentially O(1 million)
processes all actively collectively for a single ``object''. To address this
case, having some system integrated metadata services to associate names with
these object is a preferable solution.

LWFS and Sirocco have take an approach similar to the pure object stores, but
with a focus on the HPC setting. They have abandonded a fully POSIX compliant
metadata service as the default model in favor of a
container/object/fork/addrress tuple for identifying data similar to those
used for pure object stores. By having a service that addresses the object
collections that comprise a single thing, such as a variable or timestep
output, just enough metadata is maintained to make the storage system usable
without additional heavy lifting by clients.  LWFS demonstrated a POSIX-style
namespace on the side kept in sync using a transaction process like
D2T~\cite{d2t} showing that this alternative approach can support traditional
POSIX API calls even though the underlying storage system uses a different
model.

The challenge this proposal brings is in the additional metadata required to
track all of the different pieces of a single variable across the storage
hierarchy both horizontally and vertically and keeping this metadata in sync
as the storage system (Sirocco) manages data according to user requests and
system load. Further, the profiled performance for annotating data according to
future read needs also must be tracked. Overall, since the end-to-end quality
of service requirements must drive all data placement and movement decisions,
the metadata and naming service must be rich enough, and low overhead enough,
to support these operations as well as maintaining data accessibility.

In addition to the basic naming and data tracking operations, we will also need
to incorporate authorization capabilities. Sirocco currently integrates with a
Kerberos service for authentication and authorization. Given a capability
ticket, a user can access different objects as needed. This ticket structure
offers protection services typically offered on POSIX directories and files,
but can do it at the fork level instead. This allows a reduced quality data
version to be available for the general users while the high-quality version
would be limited to the data creator. This and other considerations for
security must be incorporated into the entire naming and metadat service.

The main challenge of incorporating the additional, rich metadata will be
joined by the additional challenge of coordinating with the other storage and
application-layer services to offer the best access times possible for data
stored in the system. The developed metadata services that drive data
compression and subsetting operations must have easy, consistent, and ideally
non-blocking or locking access to this metadata service. We must investigate
how to build such a metadata and naming service that also incorporates and
maintains the additional metadata required to support our advanced
functionality.
